---
title: "Homework 3_Question13"
author: "Shalini Rallapalli Venkata"
date: "February 16, 2019"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

13. In this exercise you will create some simulated data and will ???t simple linear regression models to it. Make sure to use set.seed(1) prior to starting part (a) to ensure consistent results.

(a) Using the rnorm() function, create a vector, x, containing 100 observations drawn from a N(0,1) distribution. This represents a feature,X. 
```{r}
set.seed(1)
x<-rnorm(100)
```


(b) Using the rnorm() function, create a vector, eps, containing 100 observations drawn from a N(0,0.25) distribution i.e. a normal distribution with mean zero and variance 0.25.
```{r}
eps<-rnorm(100, sd= sqrt(0.25))
```


(c) Using x and eps, generate a vector y according to the model Y =???1+0 .5X +  . (3.39) What is the length of the vector y? What are the values of ??0 and ??1 in this linear model?
```{r}
y<- -1 + 0.5*x +eps
length(y)
#length of the vector y is 100 and values of ??0 = -1 and ??1 = 0.5
```

 
 (d) Create a scatterplot displaying the relationship between x and y. Comment on what you observe. 
```{r}
plot(x,y)
#We can see that there is a linear relationship between x and y with some error by eps.
```
 
 
 (e) Fit a least squares linear model to predict y using x. Comment on the model obtained. How do ^ ??0 and ^ ??1 compare to ??0 and ??1? 
```{r}
lm.fit<- lm(y~x)
summary(lm.fit)
#??0 is -1 and ??1 hat -1.018. Similary ??1 = 1 an ??1 hat is 0.49
# We can see that the F-static value is greater than 1 thus null hypothesis can be rejected. Also R sqaure value is less but would increase with increase in number of predictors. 

```
 
 
 (f) Display the least squares line on the scatterplot obtained in (d). Draw the population regression line on the plot, in a di???erent color. Use the legend() command to create an appropriate legend. 
```{r}
plot(x,y)
abline(lm.fit, col= "green")
abline(-1, 0.5, col="blue")
legend("topleft", c("Least square", "Regression line", col= c("green", "blue")))
```
 
 
 (g) Now ???t a polynomial regression model that predicts y using x and x2. Is there evidence that the quadratic term improves the model ???t? Explain your answer. 
```{r}
lm.fit_quad<- lm(y~x+I(x^2))
summary(lm.fit_quad)
#We can see that there is a slight increase in R square and decrease in RSE. But the p-value(very high) of the x^2 term say us that there is no relationship between x^2 and y.
```
 
 
 (h) Repeat (a)-(f) after modifying the data generation process in such a way that there is less noise in the data. The model (3.39) should remain the same. You can do this by decreasing the variance of the normal distribution used to generate the error term   in (b). Describe your results. 
```{r}
set.seed(1)
x<-rnorm(100)
eps<-rnorm(100, sd= sqrt(0.125))
y<- -1 + 0.5*x +eps
plot(x,y)
fit_new<-lm(y~x)
summary(fit_new)
plot(x,y)
abline(fit_new, col= "red", lwd=2)
abline(-1, 0.5, col="blue", lwd=2)
legend("topleft", c("Least square", "Regression line"), col= c("red", "blue"))
# We can see that with the reduction in variance, the R square is increased and RSE value is decreased and the relationship is linear.

```
 
 
 (i) Repeat (a)-(f) after modifying the data generation process in such a way that there is more noise in the data. The model (3.39) should remain the same. You can do this by increasing the variance of the normal distribution used to generate the error term   in (b). Describe your results. 
```{r}
set.seed(1)
x<-rnorm(100)
eps<-rnorm(100, sd= sqrt(0.5))
y<- -1 + 0.5*x +eps
plot(x,y)
fit_new1<-lm(y~x)
summary(fit_new)
plot(x,y)
abline(fit_new1, col= "red", lwd=2)
abline(-1, 0.5, col="blue", lwd=2)
legend("topleft", c("Least square", "Regression line"), col= c("red", "blue"))
# We can see that with the reduction in variance, the R square is decreased and RSE value is increased. The relationship doesnt seems to be linear and the data points are more scattered.
```
 
 
 
 (j) What are the con???dence intervals for ??0 and ??1 based on the original data set, the noisier data set, and the less noisy data set? Comment on your results.
```{r}
confint(lm.fit)
confint(fit_new)
confint(fit_new1)
#Intrvals are centered on approximately 0.5. Lesser the noise there is more predictibility in the data

```

